---
title: "Class 12: Population Analysis"
author: "Hyeseung (Frankie) Son PID: A16025601"
format: html
editor: visual
---

# Section 1: Proportion of G/G in a population

Downloaded a CSV file from Ensemble <
https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894946-39895247;v=rs8067378;vdb=variation;vf=105535077#373531_tablePanel> 

Here we read a csv file to determine the allele frequency. 

```{r}
mxl <- read.csv("/Users/frank/Downloads/class12/373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```
```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

In the MXL population, the G|G homozygous for childhood asthma is 14%.

Now let's look at a different population. I picked GBR. 

```{r}
gbr <- read.csv("/Users/frank/Downloads/class12/373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

Find the proportion of G|G.

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100)
```
The proportion of G|G in this population is 30%, so childhood asthma is more frequent in GBR than MXL. 


